---
title: "Untitled"
author: "mrvn"
date: "2023-05-11"
output: html_document
---

```{r}
library(ggplot2)
library(readr)
library(dplyr)
```

```{r}
data <- read.csv("diabetes_binary_health_indicators_BRFSS2015.csv")
head(data)
```

```{r}
summary(data)
```

```{r}
colSums(is.na(data))
```

```{r}
dim(data)
```

```{r}
col_classes <- sapply(data, class)
col_classes
```

```{r}
ggplot(data, aes(x=BMI)) + geom_histogram()
```

```{r}
ggplot(data, aes(x=GenHlth)) + geom_histogram()
```
```{r}
ggplot(data, aes(x=MentHlth)) + geom_histogram()
```
```{r}
ggplot(data, aes(x=PhysHlth)) + geom_histogram()
```
```{r}
ggplot(data, aes(x=Age)) + geom_histogram()
```
```{r}
ggplot(data, aes(x=Income)) + geom_histogram()
```
```{r}
ggplot(data, aes(x=Education)) + geom_histogram()
```
```{r}
ggplot(data, aes(x=Diabetes_binary)) + geom_bar()
```
```{r}
# Create a new data frame with counts of each combination of diabetes_binary and highBP
counts <- data %>% group_by(Diabetes_binary, HighBP) %>% summarise(count = n())

# Create a stacked bar chart
ggplot(counts, aes(x = Diabetes_binary, y = count, fill = HighBP)) +
  geom_bar(stat = "identity")
```

```{r}
# Create a new data frame with counts of each combination of diabetes_binary and highChol
counts <- data %>% group_by(Diabetes_binary, HighChol) %>% summarise(count = n())

# Create a stacked bar chart
ggplot(counts, aes(x = Diabetes_binary, y = count, fill = HighChol)) +
  geom_bar(stat = "identity")
```

```{r}
# Create a new data frame with counts of each combination of diabetes_binary and Stroke
counts <- data %>% group_by(Diabetes_binary, Stroke) %>% summarise(count = n())

# Create a stacked bar chart
ggplot(counts, aes(x = Diabetes_binary, y = count, fill = Stroke)) +
  geom_bar(stat = "identity")
```

```{r}
# Create a new data frame with counts of each combination of diabetes_binary and Smoker
counts <- data %>% group_by(Diabetes_binary, Smoker) %>% summarise(count = n())

# Create a stacked bar chart
ggplot(counts, aes(x = Diabetes_binary, y = count, fill = Smoker)) +
  geom_bar(stat = "identity")
```

```{r}
# Create a new data frame with counts of each combination of diabetes_binary and HeartDiseaseorAttack
counts <- data %>% group_by(Diabetes_binary, HeartDiseaseorAttack) %>% summarise(count = n())

# Create a stacked bar chart
ggplot(counts, aes(x = Diabetes_binary, y = count, fill = HeartDiseaseorAttack)) +
  geom_bar(stat = "identity")
```

```{r}
data <- subset(data, !duplicated(data))
```

STATISTICAL TESTING
```{r}
# contingency table of Diabetes_binary and HighBP
table <- table(data$Diabetes_binary, data$HighBP)

# chi-squared test of independence
chisq.test(table)

```
the chi-squared value of 1480 and the very small p-value (< 2.2e-16) indicate that the observed frequencies in the table are unlikely to be due to chance alone. 
Overall, this is a good result that indicates a strong relationship between the two variables.
```{r}
# t-test of BMI between those with diabetes and those without
t.test(data$BMI ~ data$Diabetes_binary)
```
The p-value is very small, which means that the chance of observing such a large difference in BMI between the two groups by random chance alone is very low. The t-value of -99.92 suggests that the difference in BMI between the two groups is quite substantial.

```{r}

# Load the dplyr package
library(dplyr)

# Filter binary and continuous columns
binary_columns <- data %>%
  select_if(function(x) all(x %in% c(0, 1)))

continuous_columns <- data %>%
  select_if(function(x) is.numeric(x) && !all(x %in% c(0, 1)))

# Print the filtered columns
dependent = subset(binary_columns, select = c(Diabetes_binary))
binary_columns <- subset(binary_columns, select = -c(Diabetes_binary))

ordinal_columns <- subset(continuous_columns, select = c(Education, Income, Age, GenHlth))
linear_columns <- subset(continuous_columns, select = c(BMI))
continuous_columns <- subset(continuous_columns, select = -c(Education, Income, Age, GenHlth, BMI))

print(ordinal_columns)
print(linear_columns)
print(binary_columns)
print(continuous_columns)
print(dependent)

```

```{r}

print("Binary - Binary:")
for (var in colnames(binary_columns)) {
  # Binary - Binary (Chi-square and Phi coefficient)
  chi_square <- chisq.test(dependent$Diabetes_binary, binary_columns[[var]])$statistic
  table <- table(dependent$Diabetes_binary, binary_columns[[var]])
  phi_coefficient <- sqrt(chi_square / sum(table))
  
  print(paste("Column name:", var))
  print(paste("Chi-square:", chi_square))
  print(paste("Phi coefficient:", phi_coefficient))
}

```


```{r}
library(ltm)
print("Binary - Ordinal:")
for (var in colnames(ordinal_columns)) {

  correlation_result <- biserial.cor(ordinal_columns[[var]], dependent$Diabetes_binary)

  rank_biserial <- correlation_result
  
  print(paste("Column name:", var))
  print(paste("rank biserial:", rank_biserial))
}
```

```{r}
print("Binary - continuos:")
for (var in colnames(continuous_columns)) {
  print(paste("Column name:", var))
  print("point biserial:")
  print(cor.test(continuous_columns[[var]], dependent$Diabetes_binary))
}
```

```{r}
linearcor <- cor(dependent$Diabetes_binary, linear_columns$BMI, method = "pearson")
print(paste("BMI to Diabtes Binary correlation : ",linearcor))
```

```{r}
model <- glm(Diabetes_binary ~ ., data = data, family = "binomial")
```

```{r}
summary(model)
```